---
title: "2024-bean-brix-data"
format: html
---

# Setup
```{r Packages}
library(tidyverse) # for data wrangling and plotting
library(car) # for Anova function
library(lme4)
library(broom) # for model residuals extraction
library(emmeans) # for model mean extraction
library(multcomp) # for pairwise comparison letter display
library(multcompView)
```
# Import Data
```{r}
beanbrix <- read.csv("../LEM Data Stuff/data/beanbrix.csv")

beanbrix
```
# EDA tables  
```{r summary}
summary(beanbrix)
```

```{r glimpse}
glimpse(beanbrix)
```
# Wrangling
```{r beanbrixw}
beanbrixw <- beanbrix %>%
  mutate(rep = factor(rep),
         lem = factor(lem) 
         ) %>% 
  mutate(trtname = paste0(lem, "+", wincrop))

beanbrixw

summary(rcbd_dfw)

```

# EDA plots  
```{r winter crop boxplot}
ggplot(beanbrixw, aes(x = wincrop, 
                    y = avg,
                    color = wincrop)) +
  geom_boxplot() +
  geom_jitter() +
  theme(legend.position = "none")
```

```{r lem rate boxplot}
ggplot(beanbrixw, aes(x = lem, 
                    y = avg,
                    color = lem)) +
  geom_boxplot() +
  geom_jitter() +
  theme(legend.position = "none")
```
```{r nk boxplot}
ggplot(beanbrixw, aes(x = lem, 
                    y = avg,
                    color = lem)) +
  geom_boxplot() +
  geom_jitter() +
  facet_grid(.~wincrop) +
  theme(legend.position = "none")
```
# Statistical Modeling

```{r model}
# Changing to sum-to-zero contrast
options(contrasts = c("contr.sum", "contr.poly"))
beanbrixw

# Model fitting
beanbrix_mod <- lm(avg ~ rep + lem*wincrop, # rep + is functionally the only coding difference between a CRD and a RCBD design)
               data = beanbrixw)
beanbrix_mod

# Summary
summary(beanbrix_mod)
```
```{r ANOVA}
Anova(beanbrix_mod, type = 3)
```
# Linear model assumptions  
## Extracting residuals
First, let's extract our model residuals, and also create studentized residuals.  

```{r rcbd_resid}
beanbrix_resid <- augment(beanbrix_mod) %>%
  mutate(.studresid=rstudent(beanbrix_mod))

beanbrix_resid
```

## Residual independence  
- For this, we use the **fitted vs. residual plot**.  
- What we want to see: no clear pattern in residuals, random cloud of points.  
- What we do not want to see: clear pattern, for example, quadratic shape.  
- Adding a `geom_smooth()` helps with that conclusion. We want to see the smooth line and error bands comprising 0 on the y axis.  

```{r independence, homoscedasticity, and outliars}
ggplot(beanbrix_resid, aes(x=.fitted, y=.studresid))+
  geom_hline(yintercept = 0, color="red")+
  geom_point(shape = 21,
             fill = "purple", 
             size = 3,
             alpha = .7)+
  #geom_smooth()+
  geom_hline(yintercept = c(-3,3), color = "red")+
  theme_bw()
```
## Residual normality  

```{r qq}
ggplot(beanbrix_resid, aes(sample=.studresid))+
  stat_qq(  shape = 21,
            fill = "purple", 
            size = 3,
            alpha = .7
  )+
  stat_qq_line()+
  labs(x = "Theoretical quantile",
       y = "Sample quantile")+
  theme_bw()
```

```{r distribution}
ggplot(beanbrix_resid, aes(x=.studresid))+
  geom_density(color = "black",
               fill = "purple",
               alpha = .7)+
  scale_x_continuous(breaks = c(-3,0,3), limits = c(-3,3))+
  theme_bw()

```

# Model means  

```{r interaction means all}
beanbrix_means_all <- emmeans(beanbrix_mod,
                          ~lem:wincrop)

beanbrix_means_all
```
# Pairwise comparisons  
Now that we extracted means, let's perform pairwise comparisons among them.  

```{r interaction pwc all}
beanbrix_cld_all <- cld(beanbrix_means_all, 
                   reversed=T, 
                   adjust="none",
               Letters=letters,
               alpha = 0.1)

beanbrix_cld_all
```
```{r selected pwc}
beanbrix_cld_selected <- beanbrix_cld_all %>%
  as.data.frame() %>%
  mutate(letter = trimws(.group)) %>% 
  mutate(trtname = paste0(lem, "+", wincrop))
  

beanbrix_cld_selected
```
# Plot
```{r}
# New facet label names for dose variable
wincrop.labs <- c("Carrot", "Kale", "Crimson Clover")
names(wincrop.labs) <- c("C", "K", "L")
```

pp
```{r rcbd final plot}
ggplot() +
  geom_boxplot(data = beanbrixw,
               aes(x = lem,
                   y = avg,
                   fill = lem,
                   #color = trtname,
                   )
               ) +
geom_jitter(data = beanbrixw,
            aes(x = lem,
                y = avg,
                color = lem),
            alpha = 0.6) +
  geom_label(data = beanbrix_cld_selected,
            aes(x = lem,
                y = emmean,
                label = letter),
            size = 3,
            label.padding = unit(.1, "lines")) +
  labs(x = "LEM + Winter Crop",
       y = "%Brix") +
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  facet_grid(.~wincrop,
             labeller = labeller(wincrop = wincrop.labs)) +
  theme(legend.position = "none")
```



hello
